---
title: "Model Ensemble Analysis"
author: "Madeline Tucker"
date: "`r Sys.Date()`"
output: html_document
---

## Description

This script reads in five Raven model hydrograph outputs dated December 19, 2023 
and calculates:
1. Mean flow for each time series and performance statistics.
2. Apportionment period performance statistics for each time series and gauge. 
   Each period is approximately 15 days in length between April and October.
   
## Initialization

Set working directory, load libraries, and load data.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(RavenR)
library(xts)
library(plotly)
library(hydroGOF)
 
# set working directory
setwd("C:/Users/mgtucker/Documents/GitHub/SMM_Raven_hydro/")

# read in Raven hydrographs
h1 <- read_csv('calibration_members/m1/out/Hydrographs.csv')
h2 <- read_csv('calibration_members/m2/out/Hydrographs.csv')
h3 <- read_csv('calibration_members/m3/out/Hydrographs.csv')
h4<- read_csv('calibration_members/m5/out/Hydrographs.csv')
h5 <- read_csv('calibration_members/m6/out/Hydrographs.csv')

# read in naturalized flow data
# first read in csv with a summary of all gauge file names
nat_summary <- read.csv("C:/Users/mgtucker/Documents/GitHub/SMM_Raven_hydro/tables/naturalized_flows_summary.csv")
nat_summary$subbasin_id <- as.numeric(nat_summary$subbasin_id)
nat_summary$USGS_code <- as.numeric(nat_summary$USGS_code)

# loop through rvt files to read them
nat_flows <- do.call(rbind,lapply(unique(paste0('0',nat_summary$USGS_code)),function(id){
  nat_flows = rvn_rvt_read(paste0("C:/Users/mgtucker/Documents/GitHub/SMM_Raven_hydro/model/obs/naturalized/st_",id,".rvt"))
  return(nat_flows)
}))

```
   
## Mean flows for each gauge and model

Calculate mean flow for each gauge and compare with mean from data series.

```{r}
# prepare hydrographs
h_com <- rbind(h1,h2,h3,h4,h5) # combine Raven hydrographs into one data frame
h_com <- h_com[,-c(31,34)]
  
even_index <- seq(6,ncol(h_com),2)
odd_index <- seq(5,ncol(h_com),2)

h_obs <- h_com[,c(2,even_index)]
h_sim <- h_com[,c(2,odd_index)]

h_obs_avg <- h_obs %>%
  group_by(date) %>%
  summarize_all(mean,na.rm=TRUE)

# replace NaN values with NAs
for (i in 1:nrow(h_obs_avg)){
  for (j in 1:ncol(h_obs_avg)){
  h_obs_avg[i,j] <- ifelse(is.na(h_obs_avg[i,j]),NA,h_obs_avg[i,j])
  }
}

h_sim_avg <- h_sim %>%
  group_by(date) %>%
  summarize_all(mean,na.rm=TRUE)

# replace NaN values with NAs
for (i in 1:nrow(h_sim_avg)){
  for (j in 1:ncol(h_sim_avg)){
  h_sim_avg[i,j] <- ifelse(is.na(h_sim_avg[i,j]),NA,h_sim_avg[i,j])
  }
}

rm(h1,h2,h3,h4,h5)

# calculage KGE manually
  r = cor(h_obs_avg[,-1],h_sim_avg[,-1], use = 'na.or.complete') # Pearson correlation coefficient
  std_obs <- h_obs_avg[,-1] %>%
    summarize_all(sd, na.rm=TRUE) # standard deviation of observed
  std_sim <- h_sim_avg[,-1] %>%
    summarize_all(sd, na.rm=TRUE) # standard deviation of simulated
  mean_obs <- h_obs_avg[,-1] %>%
    summarize_all(mean,na.rm=TRUE) # mean of observed
  mean_sim <- h_sim_avg[,-1] %>%
    summarize_all(mean,na.rm=TRUE) # mean of simulated

KGE_ensemble <- 1 - sqrt((r-1)^2 + ((std_sim/std_obs)-1)^2 + ((mean_sim/mean_obs)-1)^2) # KGE formula

# KGE function from hydroGOF package
KGE_ensemble_test <- as.matrix(KGE(h_sim_avg[,-1],h_obs_avg[,-1])) 


```

## Apportionment Period Performance

Calculate mean observed flows across models for 15 day apportionment period as per IJC apportionment periods.
Compare with mean simulated flows for the same periods and calculate KGE.

```{r}

```


## Total Annual Outflow for Ensemble of Results

```{r}
h_obs$date <- year(h_obs$date)
obs_plot <- h_obs %>%
  group_by(date) %>%
  summarize_all(sum, na.rm=TRUE) %>%
  pivot_longer(cols = colnames(h_obs[,-1]), names_to = 'subbasin', values_to = 'flow')%>%
  mutate(type = 'observed')

h_sim$date <- year(h_sim$date)
sim_plot <- h_sim %>%
  group_by(date) %>%
  summarize_all(sum, na.rm=TRUE) %>%
  pivot_longer(cols = colnames(h_sim[,-1]), names_to = 'subbasin', values_to = 'flow')%>%
  mutate(type = 'simulated')

df_plot <- rbind(obs_plot,sim_plot)
df_plot$subbasin <- substr(df_plot$subbasin,1,9)

ggplot(df_plot)+
  geom_col(aes(x=date, y=flow, fill=type),position='dodge', width=0.75)+
  theme_bw()+ xlab('Year')+ ylab("Flow ("~m^3~"/s)") + ggtitle('Total Annual Flows')+
  theme(plot.title = element_text(hjust=0.5))+
  scale_fill_manual(name="type", values = c('#007FFF','#65CCFF'))

```
## January 18th Presentation Results

Total annual flow and time series plots in preparation for January 18, 2024 presentation to the IJC working group.

```{r}
# read in best model results dated January 17, 2024
h <- read_csv('C:/Users/mgtucker/Documents/GitHub/SMM_Raven_hydro/calibration/best_model/outputs/Hydrographs.csv')

# split dataframe into observed and simulated
even_index <- seq(6,ncol(h),2)
odd_index <- seq(5,ncol(h),2)

h_obs <- h[,c(2,even_index)]
h_sim <- h[,c(2,odd_index)]

# total annual flows plot
h_obs$date <- year(h_obs$date)
obs_plot <- h_obs %>%
  group_by(date) %>%
  summarize_all(sum, na.rm=TRUE) %>%
  pivot_longer(cols = colnames(h_obs[,-1]), names_to = 'subbasin', values_to = 'flow')%>%
  mutate(type = 'observed')

h_sim$date <- year(h_sim$date)
sim_plot <- h_sim %>%
  group_by(date) %>%
  summarize_all(sum, na.rm=TRUE) %>%
  pivot_longer(cols = colnames(h_sim[,-1]), names_to = 'subbasin', values_to = 'flow')%>%
  mutate(type = 'simulated')

df_plot <- rbind(obs_plot,sim_plot)
df_plot$subbasin <- substr(df_plot$subbasin,1,9)

ggplot(df_plot)+
  geom_col(aes(x=date, y=flow, fill=type),position='dodge', width=0.75)+
  theme_bw()+ xlab('Year')+ ylab("Flow ("~m^3~"/s)") + ggtitle('Total Annual Flows')+
  theme(plot.title = element_text(hjust=0.5))+
  scale_fill_manual(name="type", values = c('#007FFF','#65CCFF'))

# time series plot
h_obs <- h[,c(2,even_index)] # reinitialize data frames to include dates
h_sim <- h[,c(2,odd_index)]

obs_plot <- h_obs %>%
  pivot_longer(cols = colnames(h_obs[,-1]), names_to = 'subbasin', values_to = 'flow')%>%
  mutate(type = 'observed')

sim_plot <- h_sim %>%
  pivot_longer(cols = colnames(h_sim[,-1]), names_to = 'subbasin', values_to = 'flow')%>%
  mutate(type = 'simulated')

df_plot <- rbind(obs_plot,sim_plot)
df_plot$subbasin <- substr(df_plot$subbasin,1,9)

# plot St. Mary subbasins 
stmary <- c(113500,113518,113496)

df_plot_stmary <- df_plot %>%
  filter(subbasin %in% paste0('sub',stmary))

ggplot(df_plot_stmary)+
  geom_line(aes(x=date, y=flow, color=type), alpha=0.75, lwd=1)+
  theme_bw()+ xlab('Date')+ ylab("Flow ("~m^3~"/s)") + ggtitle('St. Mary Time Series by Subbasin')+
  theme(plot.title = element_text(hjust=0.5))+
  scale_color_manual(name="type", values = c('#007FFF','#FF7F00'))+
  facet_wrap(.~subbasin, scales = 'free_y', ncol=1)

# plot Milk subbasins 
milk <- c(113827,113637,114285,113991,114428,114261,114115,114119,113850,113884,114036)

df_plot_milk <- df_plot %>%
  filter(subbasin %in% paste0('sub',milk))

ggplot(df_plot_milk)+
  geom_line(aes(x=date, y=flow, color=type),alpha=0.75,lwd=1)+
  theme_bw()+ xlab('Date')+ ylab("Flow ("~m^3~"/s)") + ggtitle('Milk Time Series by Subbasin')+
  theme(plot.title = element_text(hjust=0.5))+
  scale_color_manual(name="type", values = c('#007FFF','#FF7F00'))+
  facet_wrap(.~subbasin, scales = 'free_y',ncol=3)
```


